Analysis of competitive surfing tournaments with generalized Bradley-Terry likelihoods

### 2022-03-19 inst/surfing_jsa_rebuttal.tex

and here we apply it to assess the supposed Brazilian preference for
reef-breaking waves.

Based on what hypothesis?

\textcolor{blue}{Discussed above}

Responding to ``the points system itself is intrinsically arbitrary:
the details of the points system does not affect the competitors'
behaviour but can change the overall ranking\ldots'' the referee asks:
``How could you say this, as depending on finishing spot at each
competition, the athlete is awarded a set number of points. So a
higher placing in a competition will award you a greater number of
points. Impacting upon your placing in the rankings for the subsequent
competition. And therefore, your seeding for the next competition.''

\textcolor{blue}{We agree that the original wording was confusing.
  Our point was that competitors are motivated to win, and if the
  points system awards (strictly) more points for higher ranking, then
  the competitors' behaviour is effectively driven by the ranking and
  not the points awarded.  They will be motivated to come first,
  regardless of how many points coming first is awarded (as long as
  there are more points for coming first than second etc).  }

\textcolor{blue}{
  So, for example, consider two points systems: System A awards 10
  points for a win, 7 for coming second and 1 for third; and system B
  awards 3 for a win, 2 for second and 1 for third.  Competitors will
  be equally motivated to be first, second, and third in the two
  systems, for in each system they score more points for a higher
  ranking.  But of course, if we have repeated trials (as in WSL) the
  two systems could easily result in different points-based rankings.
  We see a similar phenomenon in Formula 1 motor racing.}

\textcolor{blue}{ After discussion we believe that the manuscript
  would be improved by simply omitting the text under consideration.
  It now reads: ``\ldots the points system itself is intrinsically
  arbitrary: this would suggest that more formal statistical methods
  are needed for analysis of such datasets\ldots''.  Our emphasis was
  not on the disadvantages of the points systems used by WSL; we were
  more interested in analysing the results formally.
}

### 2022-03-19 inst/surfing_jsa.tex
referee: points system is arbitrary

Table~\ref{resultstable} shows the official results table from the
tour.  We see that Ferreira places first with 59740 points, Medina
second with 56475, and so on.  The overall ranking of the surfers is
determined by summing the season's points.  However, this system
suffers from the defect that the points system itself is intrinsically
arbitrary: the details of the points system does not affect the
competitors' behaviour but can change the overall ranking; this would
suggest that more formal statistical methods are needed for analysis
of such datasets~\citep{aldous2017}.  The Plackett-Luce likelihood
functions are not subject to such a phenomenon and furnish an
objective and coherent method for ranking competitors.  Further,
likelihood-based methods offer one advantage over points-based
systems: the ability to conduct statistically rigorous tests of
meaningful nulls.

>>>>

Table~\ref{resultstable} shows the official results table from the
tour.  We see that Ferreira places first with 59740 points, Medina
second with 56475, and so on.  The overall ranking of the surfers is
determined by summing the season's points.  However, this system
suffers from the defect that the points system itself is intrinsically
arbitrary: this would suggest that more formal statistical methods are
needed for analysis of such datasets~\citep{aldous2017}.  We suggest
here that Plackett-Luce likelihood functions furnish an objective and
coherent method for ranking competitors.  Further, likelihood-based
methods offer one advantage over points-based systems: the ability to
conduct statistically rigorous tests of meaningful nulls.
